Transcriptomic Analysis for Prognostic Value in Head and Neck Squamous Cell Carcinoma

### 2021-07-05 main.tex
Updates from Overleaf

%\begin{MyColorPar}{red}
The Cancer Genome Atlas (\acrshort{tcga}) profiled \acrshort{hnscc} (528 participants) clinical and genomic data, which were standardized and are available at a unified data portal, \acrfull{gdc} of the \acrfull{nci}.
%\cite{SeanDavis;MartinMorgan2020}
%\end{MyColorPar} % red; the rest moved into materials and methods
% to summarize the advantages of its use.
The highlight of the advantages of applying the TCGA data for cancer biomarker identification includes:
\begin{outline}
\1  To the best of our knowledge, the TCGA database is the largest collection (in both cancer types and cohort size) of comprehensive genomics with survival data available in the field of cancer research so far.  
There are many physical and social features of patients available for survival modeling. % ($X_1 ... X_n$)
The whole-genome sequencing data were harmonized across all \acrfull{gdac}s.  Many databases adopt the essential demographic data from TCGA since it has comprehensive clinical features, such as exposure to alcohol, asbestos, radioactive radon, tobacco smoking, or cigarettes.
\1  TCGA has a remarkable advantage for computational and life scientists who study cancer since useful web-based tools and API are ready to analyze and visualize TCGA data. It might be getting help quickly from the research community for trouble-shooting.
\1  Many achievements of diagnoses, treatments, and prevention with the TCGA data were already published and keep growing\cite{Tomczak2015}.\\[0.1cm]
\end{outline}

\begin{MyColorPar}{red}

@@

Xu and his colleagues\cite{Xu2021a} also used Kaplan–Meier survival curves with a log-rank test, and univariate Cox regression to screen prognostic factors.
The gene expression was used to select the preliminary biomarkers with a significant impact on patients' overall survival (\textit{P} value $< 0.01$).

@@

They validated their candidate genes by using the Gene Expression Profiling Interactive Analysis tool (GEPIA1/2 database using TCGA datasets, developed by Zefang Tang and his colleagues\cite{Tang2017a}\cite{Tang2019}

@@

\end{MyColorPar} % red

@@

The other investigators should get the rationale or revelation of the genes of interest on a specific cancer type. They should upload those genes manually onto bioinformatics tools, such as SurvExpress (http://bioinformatica.mty.itesm.mx:8080/Biomatec/ \newline
SurvivaX.jsp\cite{Aguirre-Gamboa2013}, which has been lost since Oct/2019 and currently out of funds), and analyze the cohort (e.g., \acrshort{tcga}). After downloading the survival results, they could curate plots and tables carefully.
It is not possible to scan the whole human protein-coding genome in this way.
The web-based tools might set a cutoff at the median, 1/4 quantile, or 3/4 quantile for subsequent analyses. There are several visualization software or R packages which deal with cutoff determination\cite{Abel1984}, %CRITLEVEL
such as Prognoscan\cite{Mizuno2009a}, Cutoff Finder\cite{Budczies2012}, Findcut\cite{Chang2017a}, OptimalCutpoints\cite{Cristina2019},  cutpointr (available at https://github.com/thie1e/cutpointr), and cutoffR (available at https://cran.r-project.org/web/
packages/cutoffR). 
However, non of them could perform jointly predictive of survival times and cutoff selection within subgroups of gene expression.
%  %combine the scanning of the protein-coding genes and cutoff optimization programmatically.

%%% ***
In summary, identifying predictive biomarkers for selecting standard-of-care or advanced systemic therapy\cite{Cristina2019} in \acrshort{hnscc} is crucial.
Our approach describes an in-house workflow implemented in the R script, which runs on the Rstudio server.
Its function includes data retrieving and pre-processing, feature selection, sliding-window cutoff mining, Kaplan-Meier survival analysis, Cox proportional hazard modeling, and biomarker discovery.
The independent HNSCC dataset (GSE65858) was used to validate this strategy.
The workflow, shown in Figure \ref{fig:figure1}, scanned 20,500 human protein-coding genes of the \acrshort{tcga} \acrshort{hnscc} cohort to yield a model with biomarker estimate.

@@

\begin{MyColorPar}{red}
TCGA HNSCC cohort was applied for exploration of the biomarker candidate.
A total of 9416 Kaplan-Meier plots (under sliding-window cutoff selection) with associated Cox's univariate and multivariate tables were generated by Cox modeling (see Figure ~\ref{fig:figure1}) and justified by the ranking of hazard ratios.
%By uncorrected P-value below 0.05, we selected 967 genes in which \acrfull{hr} is greater than 1.5 or less than 0.5 (see Figure \ref{fig:figure2}(a) univariate, and Figure \ref{fig:figure2}(b)  multivariate plots). 
The 967 out of 9416 genes were kept by criteria of Kaplan-Meier \textit{P} value ($< 0.05$) and \acrfull{hr} derived from Cox's model (please see Figure \ref{fig:figure2}(a), (b), initial trial).
In the next step, a Bonferroni P-value correction was used to yield 20 genes, under the stringent criteria (see Figure \ref{fig:figure2}(c), (d)).

@@

% [2021/06/13-17] GSE65858
%CAMK2N1, IL19, and FCGBP is validated by
%candidate_fdr100
%PvalueTex consensus with (x) GSE2837 to yield 3 candidates
In the study of GSE65858 cohort with median cutoffs, CAMK2N1, CALML5, and FCGBP (3 out of those 20 genes discovered in the TCGA cohort), keep ahead of the curve by their FDR-corrected \textit{P} value ($< 0.05$), and Cox's HR ($> 1.8$ or $<0.6$) (please see Table \ref{tab:table_top3_GSE65858}).
% (ok) (FDR KM \textit{P} value) <= # only MSMB KM FDR-adjusted P value = 0.002
% Consensus 3 genes in Kaplan-Meier survival and Cox's model (dataset: GSE65858)
%Gene.Symbol KM.Pvalue   FDR.KMpvalue    uni_HR  FDR.Pvalue(in TCGA)
%CAMK2N1 0.00687 0.03782500  1.814   1.628308e-05
%FCGBP   0.01020 0.03886709  0.573   4.827833e-05
%IL19    0.03390   0.04738901    0.630   6.543871e-06
%IL19    0.00269  0.03068073 0.630  6.543871e-06
%candidate_fdr100
However, the significance of the other 17 genes are replaced by genes such as DUSP6, MSMB, and RBM11 (please see Figure \ref{fig:hazards534}).
%This GSE65858 cohort has been applied for selection to generate our candidate genes: \textcolor{red}{CAMK2N1}, \textcolor{red}{IL19} (two probes), and , \textcolor{red}{FCGBP}.
Conversly, there are 22 genes, which have greater hazard ratio ($> 1.8$ or $< 0.6$) in GSE65858 cohort (Figure \ref{fig:hazards534}), drop their hazard ratio between 0.6 and 1.5 in the study of TCGA HNSCC (Figure \ref{fig:hazards3}).
Thus, there is a consensus between the TCGA and GSE65858 cohorts that CAMK2N1, CALML5, and FCGBP are significant candidates for the HNSCC biomarker.
%Finally, these 3 candidates have also been confirmed by a HNSCC study using the GSE2837 dataset. Please see the Kaplan-Meier plots in Supplementary Figure S2.%\ref{fig:fig_GSE2837}.
\end{MyColorPar} % end red

@@

\begin{MyColorPar}{red}

@@

We explain aforementioned situation by examples.
%***NDFIP1  % 給他們一個機會
Reviewing the special case of genes, such as NDFIP1, DKC1, PNMA5, and NPB,
we noticed that NDFIP1, with a \textit{P} value of 0.05 at 50\% quantile (median) cutoff, could achieve a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff.
NDFIP1 has the FDR-adjusted \textit{P} value as \num[round-precision=3, round-mode=figures,
scientific-notation=true]{0.0001074233} (please see supplementary figure S1, a "S"- or "W"-shaped acute-bending on the \textit{P} value plot). % it is \ref{fig:pvaluePlot_NDFIP1
%Moreover, the gene NDFIP1, one of our 20 preliminary candidates, has a \textit{P} value (around $0.05$) at 50\% quantile cutoff, achieving a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff (please see Figure \ref{fig:pvaluePlot_NDFIP1}, and also in supplementary figure S1).
%After the FDR correction, NDFIP1 still has \textit{P} value of \num[round-precision=3, round-mode=figures, scientific-notation=true]{0.0001074233}.
% the effect size
However, it was excluded as a candidate by small effect size (HR = 1.33 in GSE65858) less than 1.8.
%%%  and HR = 1.42 in GSE2837

The other example is IL19. It has a \textit{P} value plot with acute S-curve bending at the median zone, that lets the FDR-adjusted \textit{P} value has large difference between 50\% quantile cutoff (FDR 0.115) and optimal 48\% quantile cutoff (FDR \num{6.54e-6}). This optimal cutoff method could boost it's statistic significance to pass the correction by both FDR and Bonferroni methods. Even IL19 has competed as a candidate by its effect size (HR = 0.472 in TCGA cohort), but failed by the validation with GSE65858 cohort (HR = 0.630 and FDR-KM \textit{P} = 0.031).

\end{MyColorPar} % red

% unclean data: moved from introduction to discussion 
\subsection{Technical Consideration}
\begin{MyColorPar}{red} %%% ***
There are two essential points of biomarker discovery from survival analysis of the TCGA HNSCC dataset.

First, although \acrshort{tcga} genomics data were harmonized, 
% three things were found during our investigating and data cleaning:
the pre-processing of TCGA RNA-Seq in our workflow has been done as follows: 
\begin{outline}
\1 HNSCC samples without complete clinical information have been removed;
\1 Null expressed genes in more than 30\% of the HNSCC samples have been excluded; % or 50\%
\1 Updated number of protein-coding genes in the TCGA HNSCC is 20500.
\end{outline}
After investigation of mRNA expression dataset obtained through NCI's Firehose API, we found that expression value of two genes (gene ID: 9906 and gene ID: 728661) was saved together under the entity of gene symbol "SLC35E2".
The expression file of SLC35E2 has almost double in size than those of SLC35E1 or SLC35E3.
% http://gepia2.cancer-pku.cn/#survival is SLC35E2; 證明 TCGA has error
% [line 1549]
%HNSCC.mRNA.Exp.SLC35E2.Fire.Rda
%[line 1654]
%for gene correction (20499 to 20500) from SLC35E2 to 2 genes:
%NCBI Reference Sequences (RefSeq) National Center for Biotechnology Information Reference Sequence (NCBI) RefSeq Release 38 (November 7, 2009) 
According to the Human Gene Database (available at https://www.genecards.org/Search/Keyword?\\queryString=SLC35E2), SLC35E2A (Gene ID: 9906) and SLC35E2B (Gene ID: 728661) should be the correct entities for the TCGA HNSCC dataset.
% https://www.ncbi.nlm.nih.gov/gene/?term=(SLC35E2)+AND+%22Homo+sapiens%22
% keyword: (SLC35E2) AND "Homo sapiens" 
SLC35E2 is the previous symbol of SLC35E2A (reference at https://www.genenames.org/data/gene-symbol-report/\#!/hgnc\_id/HGNC:20863).
Thus, we made reassignment of the expression value of SLC35E2A and SLC35E2B and updated the number of protein-coding genes in this TCGA HNSCC dataset from 20499 to 20500.
% updated on 18-May-2021
%\end{outline}

@@

\1 32.2\% event has "one group" issue in confusion matrix of Chi-square test in Cox regression;
\1 21.05\% error occurs by "one group" issue in log-rank test (survdiff or survdiff.fit function in R package "survival") in Kaplan-Meier estimate;
%1) ok [1] "[ 4 / 20083 ] Generating median-cut KM plot of A1BG"
%Error in survdiff.fit(y, groups, strata.keep, rho) : at A1CF
% error at 16115 gene ""; why? 518 values all are -0.04203 at T
%  There is only 1 group 
%This is because of skewed distribution of the expression value. 
% 67/518 =  13% cases >1, while 450 (87%) cases = -0.30949; so median is -0.30949; 
% all cases belong to >= median => 改為 > median 即可
% 2) % survfit 16114 has data set has no non-missing observations: NA's 113/518 cases
% median = NaN :-); 
% solved by cutoff <- median(gset_OS_mRNA2[, i], na.rm = TRUE)
% and NaN imputation as median
\1 0.78\% has unknown reason (so those 159 genes has been excluded in our workflow).
\end{outline}
These technical problems could not be detected prior to program running.
It might be due to skewed distribution of the expression value or even random error in 20500 RNA-Seq data. % heterogeneity

\end{MyColorPar} %%% red

@@

\begin{MyColorPar}{red}
There are 81 physical, pathological and social conditions derived from participants available for survival modeling in the TCGA, % since it has comprehensive clinical features, 
such as age, gender, residual tumor, vital status, days-to-last-followup, cancer stage, smoking duration, exposure to alcohol, asbestos, radioactive radon. 
However, the TCGA did not collect other features related to holistic care.
% Input $X_1...X_n$ (e.g., patients features: such as age, gender, gene expression, cancer stage; emotional 情緒(心)
Going for holistic cancer care\cite{Mehta2019}\cite{Iftikhar2021}, spiritual and emotional condition is equally essential comparing with physical and social status (see Figure \ref{fig:figure5}).
Psychosocial stress is associated with cancer incidence\cite{Lutgendorf2010}\cite{Powell2013}\cite{Iftikhar2021}, metastasis\cite{Lutgendorf2010}\cite{Moreno-Smith2010}\cite{Du2020}\cite{Xu2021}, and poor survival\cite{Chida2008}.
These impacts might be mediated through the hypothalamic-pituitary-adrenal (HPA) axis\cite{Hsiao2012}.
Holistic healthcare providers will engage his/her patients with eye-to-eye contact in terms of mind-to-mind connection. Their empathy, sympathy, and compassion are induced by the suffering of patients from those diseases. They will try to treat patients by prescribing medicine (and "themselves") or performing surgery.
Thus, the healing resilience of patients will be induced by unconditional positive regard. % of healing.
The patients will trust those who take care of them and have the confidence to increase the capacity to recover from diseases through a mind-brain-body connection manner (see Figure \ref{fig:figure5}).
\end{MyColorPar} % red

@@

%\label{key:cutoff}
To evaluate the effect of gene expression on the patient's survival, we introduce a sliding window cutoff selection by stratifying patients with Kaplan-Meier survival analysis according to each gene's low/high expression.
Our cutofFinder\_func subroutine employs the minimum \textit{P} value approach to recognizing cutoff points in continuous gene expression measurement for patients sub-population.
First, patients were ordered by \acrshort{rnaseq} value (RSEM) of a given gene. Next, patients were stratified at a serial cut (counted by person ranked in 30\% to 70\% percentile of the cohort; please see Figure \ref{fig:figure1} cutoff engine). The survival risk differences of the two groups were estimated by log-rank test to yield around 165 Kaplan-Meier \textit{P} values for each gene.
Then, the optimal cutoff of \acrshort{rnaseq}, giving the minimum \textit{P} value, was selected by the cutofFinder\_func subroutine.
This iteration method could calculate all possible cutoff of each gene expression in this cohort. At each run of cutofFinder\_func function call for an individual gene, it returned an optimal cutoff at specific patient grouping  (e.g., low-expression in 262 persons versus high-expression in 152 persons with gene \acrlong{CAMK2N1}, \acrshort{CAMK2N1}, please see Figure \ref{fig:figure4}). The cutoff would be returned to the main program to allow downstream Cox survival analysis. The percentile range we applied as 30\% to 70\% was used to avoid a small grouping effect\cite{Miller1982}\cite{Mizuno2009a}.
In case there was no significant \textit{P} value, a median expression of this gene was set as its cutpoint as usual.
\begin{MyColorPar}{red}
% (stored at HNSCC_survivalAnalysis_marginS_[gene].Rda; file20500_list.txt)
The \acrfull{fdr} ($< 0.05$) correction\cite{Benjamini1995a} shows which genes should be retained for subsequent univariate and multivariate analysis.
% of Cox modeling
It ensures the control of type I error of multiple test \textit{P} values during our cutoff finding procedure.
Then Bonferroni adjustment of that \textit{P} values has been used for candidate selection.
\end{MyColorPar} % red

@@

% reviewer comment; Answer 2-4 (Cox regression)
\begin{MyColorPar}{red}
%The Cox proportional-hazards regression model\cite{Cox1972}\cite{Andersen1982} is commonly used for modeling survival analysis data. It allows analyzing survival for one or more predictor variables and to provides the effect size (coefficients) for them\cite{Bradburn2003b}. % Classification trees and artificial neural networks in survival analysis
The Cox model is also accounting for confounding factors\cite{Magen2019}.
It is expressed by the hazard function denoted by h(t). 
The hazard function represents the risk of a specific event (e.g., death) at time t. It can be estimated as follow:

@@

In a multivariate test, those covariates used include the clinicopathological features (gender, age at diagnosed separated by 65 year-old, clinical tumor size [T1-2/T3-4], clinical nodal status [N0/N+], clinical distant metastasis [M0/M1], TNM staging [stage 1-2/stage 3-4], surgical margin status [negative/positive] and tobacco exposure [low/high]), and gene expression level [low/high] defined by the cutoff.
All covariates have been pooled in the hazard function h(t) to estimate their impact on the overall survival.
\end{MyColorPar} % red
%%

>>>>

%\begin{MyColorPar}{red}
The Cancer Genome Atlas (\acrshort{tcga}) profiled \acrshort{hnscc} (528 participants) clinical and genomic data, which were standardized and are available at a unified data portal, \acrfull{gdc} of the \acrfull{nci}.
%\cite{SeanDavis;MartinMorgan2020}
%\end{MyColorPar} % red; the rest moved into materials and methods
% to summarize the advantages of its use.
The highlight of the advantages of applying the TCGA data for cancer biomarker identification includes:
\begin{outline}
\1  To the best of our knowledge, the TCGA database is the largest collection (in both cancer types and cohort size, especially in HNSCC) of comprehensive genomics with survival data available in the field of cancer research so far.
%There are many physical and social features of patients available for survival modeling. % ($X_1 ... X_n$)
The whole-genome sequencing data were harmonized across all \acrfull{gdac}s.  Many databases adopt the essential demographic data from TCGA since it has comprehensive physical and social features of patients, such as exposure to alcohol, asbestos, radioactive radon, tobacco smoking, or cigarettes.
\1  TCGA has a remarkable advantage for computational and life scientists who study cancer since useful web-based tools and \acrfull{api} are ready to analyze and visualize TCGA data. It might be getting help quickly from the research community for trouble-shooting.
\1  Many achievements of diagnoses, treatments, and prevention from the TCGA data were already published and keep growing\cite{Tomczak2015}.\\[0.1cm]
\end{outline}

%\begin{MyColorPar}{red}

@@

Xu and his colleagues\cite{Xu2021a} also identified their biomarkers by  \acrshort{degs} analysis.
%The differential expression of gene was used to select the preliminary biomarkers.
The significant impact of genes on patients' overall survival was evaluated by Kaplan–Meier survival curves with a log-rank test (\textit{P} value $< 0.01$), and univariate Cox regression.

@@

They validated the candidate genes by using the web-based Tools of Gene Expression Profiling Interactive Analysis tool---GEPIA and Human protein atlas (HPA) databases.
GEPIA was developed, using TCGA datasets, by Zefang Tang and his colleagues (version 1\cite{Tang2017a}, version 2\cite{Tang2019}, and GEPIA2021\cite{Li2021}, available at http://gepia2021.cancer-pku.cn/).

@@

%\end{MyColorPar} % red

@@

%http://bioinformatica.mty.itesm.mx:8080/Biomatec/SurvivaX.jsp
% bioinformatics
The other investigators should get the rationale or revelation of the genes of interest on a specific cancer type. They should upload those genes manually onto web-based tools, such as SurvExpress\cite{Aguirre-Gamboa2013} (available at http://bioinformatica.mty.itesm.mx:8080/\\Biomatec/SurvivaX.jsp), then analyze the cohort of interest (e.g., \acrshort{tcga}). After downloading the survival results, they could curate plots and tables carefully.
It is not possible to scan the whole human protein-coding genome in this way.
The web-based tools might set a cutoff at the median, 1/4 quantile, or 3/4 quantile for subsequent analyses. There are several visualization software or R packages which deal with cutoff determination\cite{Abel1984}, %CRITLEVEL
such as Prognoscan\cite{Mizuno2009a}, Cutoff Finder\cite{Budczies2012}, Findcut\cite{Chang2017a}, OptimalCutpoints\cite{Cristina2019},  cutpointr (available at https://github.com/thie1e/cutpointr), and cutoffR (available at https://cran.r-project.org/web/packages/cutoffR). 
However, non of them could perform jointly survival analysis with cutoff selection and whole-genome scanning.
%  %combine the scanning of the protein-coding genes and cutoff optimization programmatically.

%%% ***
In summary, identifying predictive biomarkers for selecting standard-of-care or advanced systemic therapy\cite{Cristina2019} in \acrshort{hnscc} is crucial.
Our approach describes an in-house workflow implemented in the R script, which runs on the Rstudio server.
Its function includes data retrieving and pre-processing, feature selection, sliding-window cutoff selection, Kaplan-Meier survival analysis, Cox proportional hazard modeling, and biomarker discovery.
The independent HNSCC dataset (GSE65858)\cite{Wichmann2015} was used to validate this strategy.
The workflow, shown in Figure \ref{fig:figure1}, has scanned 20,500 human protein-coding genes of the \acrshort{tcga} \acrshort{hnscc} cohort to yield a model with biomarker estimate by gene-expression based survival analysis.

@@

%\begin{MyColorPar}{red}
TCGA HNSCC cohort was applied for exploration of the biomarker candidate.
A total of 9416 Kaplan-Meier plots (under sliding-window cutoff selection) with associated Cox's univariate and multivariate tables were generated by Cox modeling (see Figure ~\ref{fig:figure1}) and justified by the ranking of hazard ratios.
%By uncorrected P-value below 0.05, we selected 967 genes in which \acrfull{hr} is greater than 1.5 or less than 0.5 (see Figure \ref{fig:figure2}(a) univariate, and Figure \ref{fig:figure2}(b)  multivariate plots). 
The 967 out of 9416 genes were kept by criteria of Kaplan-Meier \textit{P} value ($< 0.05$) and \acrfull{hr} derived from Cox's model (please see Figure \ref{fig:figure2}(a), (b), initial trial).
In the next step, a Bonferroni P-value correction was used to yield 20 genes, under the stringent criteria (see Figure \ref{fig:figure2}(c), (d)).

@@

% [2021/06/13-17] GSE65858
%CAMK2N1, IL19, and FCGBP is validated by
%candidate_fdr100
%PvalueTex consensus with (x) GSE2837 to yield 3 candidates
In the study of GSE65858 cohort with median cutoffs, CAMK2N1, CALML5, and FCGBP (3 out of those 20 genes discovered in the TCGA cohort), keep ahead of the curve by their FDR-corrected \textit{P} value ($< 0.05$), and Cox's HR ($> 1.8$ or $<0.6$) (please see Table \ref{tab:table_top3_GSE65858}).
% (ok) (FDR KM \textit{P} value) <= # only MSMB KM FDR-adjusted P value = 0.002
% Consensus 3 genes in Kaplan-Meier survival and Cox's model (dataset: GSE65858)
%Gene.Symbol KM.Pvalue   FDR.KMpvalue    uni_HR  FDR.Pvalue(in TCGA)
%CAMK2N1 0.00687 0.03782500  1.814   1.628308e-05
%FCGBP   0.01020 0.03886709  0.573   4.827833e-05
%IL19    0.03390   0.04738901    0.630   6.543871e-06
%IL19    0.00269  0.03068073 0.630  6.543871e-06
%candidate_fdr100
However, the significance of the other 17 genes are replaced by genes such as DUSP6, MSMB, and RBM11 (please see Figure \ref{fig:hazards534}).
%This GSE65858 cohort has been applied for selection to generate our candidate genes: \textcolor{red}{CAMK2N1}, \textcolor{red}{IL19} (two probes), and , \textcolor{red}{FCGBP}.
Conversly, there are 22 genes, which have greater hazard ratio ($> 1.8$ or $< 0.6$) in GSE65858 cohort (Figure \ref{fig:hazards534}), drop their hazard ratio between 0.6 and 1.5 in the study of TCGA HNSCC (Figure \ref{fig:hazards3}).
Thus, there is a consensus between the TCGA and GSE65858 cohorts that CAMK2N1, CALML5, and FCGBP are significant candidates for the HNSCC biomarker.
%Finally, these 3 candidates have also been confirmed by a HNSCC study using the GSE2837 dataset. Please see the Kaplan-Meier plots in Supplementary Figure S2.%\ref{fig:fig_GSE2837}.
%\end{MyColorPar} % end red

@@

%\begin{MyColorPar}{red}

@@

We explain aforementioned situation by examples.
%***NDFIP1  % 給他們一個機會
Reviewing the special case of genes, such as NDFIP1, DKC1, PNMA5, and NPB,
we noticed that NDFIP1, with a \textit{P} value of 0.05 at 50\% quantile (median) cutoff, could achieve a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff.
NDFIP1 has the FDR-adjusted \textit{P} value as \num[round-precision=3, round-mode=figures,
scientific-notation=true]{0.0001074233} (please see supplementary Figure S1, a "S"- or "W"-shaped acute-bending on the \textit{P} value plot). % it is \ref{fig:pvaluePlot_NDFIP1
%Moreover, the gene NDFIP1, one of our 20 preliminary candidates, has a \textit{P} value (around $0.05$) at 50\% quantile cutoff, achieving a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff (please see Figure \ref{fig:pvaluePlot_NDFIP1}, and also in supplementary figure S1).
%After the FDR correction, NDFIP1 still has \textit{P} value of \num[round-precision=3, round-mode=figures, scientific-notation=true]{0.0001074233}.
% the effect size
However, it was excluded as a candidate by small effect size (HR = 1.33 in GSE65858) less than 1.8.
%%%  and HR = 1.42 in GSE2837

The other example is IL19. It has a \textit{P} value plot with acute S-curve bending at the median zone, that lets the FDR-adjusted \textit{P} value has large difference between 50\% quantile cutoff (FDR 0.115) and optimal 48\% quantile cutoff (FDR \num{6.54e-6}). This optimal cutoff method could boost it's statistic significance to pass the correction by both FDR and Bonferroni methods. Even IL19 has competed as a candidate by its effect size (HR = 0.472 in TCGA cohort), but failed by the validation with GSE65858 cohort (small effect size as HR = 0.630, and FDR-KM \textit{P} = 0.031).

%\end{MyColorPar} % red

% unclean data: moved from introduction to discussion 
\subsection{Technical Consideration}
%\begin{MyColorPar}{red} %%% ***
There are two essential points of biomarker discovery from survival analysis of the TCGA HNSCC dataset.

First, since \acrshort{tcga} genomic data were harmonized, 
% three things were found during our investigating and data cleaning:
the pre-processing of TCGA RNA-Seq in our workflow has been done as follows: 
\begin{outline}
\1 HNSCC samples without complete clinical information have been removed;
\1 Null expressed genes in more than 30\% of the HNSCC samples have been excluded; % or 50\%
\1 Updated number of protein-coding genes in the TCGA HNSCC is 20500.
\end{outline}
After investigation of mRNA expression dataset obtained through NCI's Firehose API, we found that expression value of two genes (gene ID: 9906 and gene ID: 728661) was saved together under the entity of gene symbol "SLC35E2".
The expression file of SLC35E2 has almost double in size than those of SLC35E1 or SLC35E3.
% http://gepia2.cancer-pku.cn/#survival is SLC35E2; 證明 TCGA has error
% [line 1549]
%HNSCC.mRNA.Exp.SLC35E2.Fire.Rda
%[line 1654]
%for gene correction (20499 to 20500) from SLC35E2 to 2 genes:
%NCBI Reference Sequences (RefSeq) National Center for Biotechnology Information Reference Sequence (NCBI) RefSeq Release 38 (November 7, 2009) 
According to the Human Gene Database (available at https://www.genecards.org/Search/Keyword?\\queryString=SLC35E2), SLC35E2A (Gene ID: 9906) and SLC35E2B (Gene ID: 728661) should be the correct entities for the TCGA HNSCC dataset.
% https://www.ncbi.nlm.nih.gov/gene/?term=(SLC35E2)+AND+%22Homo+sapiens%22
% keyword: (SLC35E2) AND "Homo sapiens" 
SLC35E2 is the previous symbol of SLC35E2A (reference at https://www.genenames.org/data/gene-symbol-report/\#!/hgnc\_id/HGNC:20863).
Thus, we made reassignment of the expression value of SLC35E2A and SLC35E2B and updated the number of protein-coding genes in this TCGA HNSCC dataset from 20499 to 20500.
% updated on 18-May-2021
%\end{outline}

@@

\1 32.2\% event has "one group" issue in confusion matrix of Chi-square test in Cox regression;
\1 21.05\% error occurs by "one group" issue in log-rank test (survdiff or survdiff.fit function in R package "survival") in Kaplan-Meier estimate;
%1) ok [1] "[ 4 / 20083 ] Generating median-cut KM plot of A1BG"
%Error in survdiff.fit(y, groups, strata.keep, rho) : at A1CF
% error at 16115 gene ""; why? 518 values all are -0.04203 at T
%  There is only 1 group 
%This is because of skewed distribution of the expression value. 
% 67/518 =  13% cases >1, while 450 (87%) cases = -0.30949; so median is -0.30949; 
% all cases belong to >= median => 改為 > median 即可
% 2) % survfit 16114 has data set has no non-missing observations: NA's 113/518 cases
% median = NaN :-); 
% solved by cutoff <- median(gset_OS_mRNA2[, i], na.rm = TRUE)
% and NaN imputation as median
\1 0.78\% has unknown reason (so those 159 genes has been excluded in our workflow).
\end{outline}
These technical problems could not be detected prior to program running.
It might be due to skewed distribution of the expression value or even random error derived during the RNA sequencing procedure. % heterogeneity

%\end{MyColorPar} %%% red

@@

%\begin{MyColorPar}{red}
There are 81 physical, pathological and social conditions derived from participants available for survival modeling in the TCGA, % since it has comprehensive clinical features, 
such as age, gender, residual tumor, vital status, days-to-last-followup, cancer stage, smoking duration, exposure to alcohol, asbestos, radioactive radon. 
However, the TCGA did not collect other features related to holistic care.
% Input $X_1...X_n$ (e.g., patients features: such as age, gender, gene expression, cancer stage; emotional 情緒(心)
Going for holistic cancer care\cite{Mehta2019}\cite{Iftikhar2021}, spiritual and emotional condition is equally essential comparing with physical and social status (see Figure \ref{fig:figure5}).
Psychosocial stress is associated with cancer incidence\cite{Lutgendorf2010}\cite{Powell2013}\cite{Iftikhar2021}, metastasis\cite{Lutgendorf2010}\cite{Moreno-Smith2010}\cite{Du2020}\cite{Xu2021}, and poor survival\cite{Chida2008}.
These impacts might be mediated through the hypothalamic-pituitary-adrenal (HPA) axis\cite{Hsiao2012}.
Holistic healthcare providers will engage his/her patients with eye-to-eye contact in terms of mind-to-mind connection. Their empathy, sympathy, and compassion are induced by the suffering of patients from those diseases. They will try to treat patients by prescribing medicine (and "themselves") or performing surgery.
Thus, the healing resilience of patients will be induced by unconditional positive regard. % of healing.
The patients will trust those who take care of them and have the confidence to increase the capacity to recover from diseases through a mind-brain-body connection manner (see Figure \ref{fig:figure5}).
%\end{MyColorPar} % red

@@

%\label{key:cutoff}
To evaluate the effect of gene expression on the patient's survival, we introduce a sliding window cutoff selection by stratifying patients with Kaplan-Meier survival analysis according to each gene's low/high expression.
Our cutofFinder\_func subroutine employs the minimum \textit{P} value approach to recognizing cutoff points in continuous gene expression measurement for patients sub-population.
First, patients were ordered by \acrshort{rnaseq} value (RSEM) of a given gene. Next, patients were stratified at a serial cut (counted by person ranked in 30\% to 70\% percentile of the cohort; please see Figure \ref{fig:figure1} cutoff engine). The survival risk differences of the two groups were estimated by log-rank test to yield around 165 Kaplan-Meier \textit{P} values for each gene.
Then, the optimal cutoff of \acrshort{rnaseq}, giving the minimum \textit{P} value, was selected by the cutofFinder\_func subroutine.
This iteration method could calculate all possible cutoff of each gene expression in this cohort. At each run of cutofFinder\_func function call for an individual gene, it returned an optimal cutoff at specific patient grouping  (e.g., low-expression in 262 persons versus high-expression in 152 persons with gene \acrlong{CAMK2N1}, \acrshort{CAMK2N1}, please see Figure \ref{fig:figure4}). The cutoff would be returned to the main program to allow downstream Cox survival analysis. The percentile range we applied as 30\% to 70\% was used to avoid a small grouping effect\cite{Miller1982}\cite{Mizuno2009a}.
In case there was no significant \textit{P} value, a median expression of this gene was set as its cutpoint as usual.
%\begin{MyColorPar}{red}
% (stored at HNSCC_survivalAnalysis_marginS_[gene].Rda; file20500_list.txt)
The \acrfull{fdr} ($< 0.05$) correction\cite{Benjamini1995a} shows which genes should be retained for subsequent univariate and multivariate analysis.
% of Cox modeling
It ensures the control of type I error of multiple test \textit{P} values during our cutoff finding procedure.
Then Bonferroni adjustment of that \textit{P} values has been used for candidate selection.
%\end{MyColorPar} % red

@@

% reviewer comment; Answer 2-4 (Cox regression)
%\begin{MyColorPar}{red}
The Cox proportional-hazards regression model\cite{Cox1972}\cite{Andersen1982} is commonly used for modeling survival data. It allows analyzing survival for one or more variables and to provides the effect size (coefficients, i.e., hazard ratios) for them\cite{Bradburn2003b}. % Classification trees and artificial neural networks in survival analysis
The Cox model is also accounting for confounding factors\cite{Magen2019}.
It is expressed by the hazard function denoted by h(t). 
The hazard function represents the risk of a specific event (e.g., death) at time t. It can be estimated as follow:

@@

In a multivariate test, those covariates used include the clinicopathological features (gender, age at diagnosed separated by 65 year-old, clinical tumor size [T1-2/T3-4], clinical nodal status [N0/N+], clinical distant metastasis [M0/M1], TNM staging [stage 1-2/stage 3-4], surgical margin status [negative/positive] and tobacco exposure [low/high]), and gene expression level [low/high] defined by the cutoff.
All covariates have been pooled in the hazard function h(t) to estimate their impact on the overall survival.
%\end{MyColorPar} % red
%%

### 2021-07-05 main_redMarked.tex
Updates from Overleaf

% reviewer comment; Answer 2-4 (Cox regression)
\begin{MyColorPar}{red}
%The Cox proportional-hazards regression model\cite{Cox1972}\cite{Andersen1982} is commonly used for modeling survival analysis data. It allows analyzing survival for one or more predictor variables and to provides the effect size (coefficients) for them\cite{Bradburn2003b}. % Classification trees and artificial neural networks in survival analysis
The Cox model is also accounting for confounding factors\cite{Magen2019}.
It is expressed by the hazard function denoted by h(t). 
The hazard function represents the risk of a specific event (e.g., death) at time t. It can be estimated as follow:

>>>>

% reviewer comment; Answer 2-4 (Cox regression)
\begin{MyColorPar}{red}
The Cox proportional-hazards regression model\cite{Cox1972}\cite{Andersen1982} is commonly used for modeling survival data. It allows analyzing survival for one or more variables and to provides the effect size (coefficients, i.e., hazard ratios) for them\cite{Bradburn2003b}. % Classification trees and artificial neural networks in survival analysis
The Cox model is also accounting for confounding factors\cite{Magen2019}.
It is expressed by the hazard function denoted by h(t). 
The hazard function represents the risk of a specific event (e.g., death) at time t. It can be estimated as follow:
